\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath,amscd}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{caption}
\usepackage{floatpag}
\usepackage[english]{babel}
\linespread{1}

\newtheorem{theo}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}[theo]{Corollary}
\newtheorem{defn}{Definition}
\newtheorem{con}{Conjecture}

\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Real}{Re}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

% Title
\title{$1^{st}$ year PhD report}
\author{Mauro D'Arcangelo \\ Supervisor: Prof. J. W. Barrett}
\date{}

\begin{document}

\maketitle


% Table of contents
\tableofcontents

% Sections
\section{Introduction}\label{intro}
\input{intro_bare}

\section{Results}\label{results}
\input{HMC_bare}
\input{sven_bare}
\input{IKKT_bare}

\section{Future directions}\label{future}
\input{spvar_bare}
\input{matter_bare}
\input{RHMC_bare}

\input{appendix_bare}

\begin{thebibliography}{99} % alternatively use BibTeX

\bibitem{barrettLOR}
J.~W. Barrett,
``Lorentzian version of the noncommutative geometry of the standard model of particle physics'',
{\it J. Math.\ Phys.\ \bf 48}, 012303 (2007).

\bibitem{barrett}
J.~W. Barrett,
``Matrix geometries and fuzzy spaces as finite spectral triples'',
{\it J. Math.\ Phys.\ \bf 56}, 082301 (2015).

\bibitem{barrettglaser}
J.~W. Barrett and L. Glaser,
``Monte Carlo simulations of random non-commutative geometries'',
{\it J. Phys.\ A\ \bf 49}, 24, 245001 (2016).

\bibitem{betan}
M. Betancourt,
``A Conceptual Introduction to Hamiltonian Monte Carlo'',
(arXiv:1701.02434v1, 2017).

\bibitem{bhat}
U. Bhat, M. Shrestha and L. H\'ebert-Dufresne,
``Exotic phase transitions of k-cores in clustered networks'',
{\it Phys.\ Rev. \ E \ \bf 95}, 012314 (2017).

\bibitem{bittner}
E. Bittner and W. Janke,
``Parallel-tempering cluster algorithm for computer simulations of critical phenomena'',
{\it Phys.\ Rev. \ E \ \bf 84}, 036701 (2011).

\bibitem{connesSAP}
A.~H. Chamseddine and A. Connes,
``The spectral action principle'',
{\it Commun.\ Math.\ Phys. \ \bf 186}, 731--750 (1997).

\bibitem{connesNEUT}
A.~H. Chamseddine, A. Connes and M. Marcolli,
``Gravity and the standard model with neutrino mixing'',
{\it Adv.\ Theor. \ Math.\ Phys. \ \bf 11}, 991--1089 (2007).

\bibitem{pol}
P.  Colomer-de-Sim\'on and M. Bogu\~n\'a,
``Double Percolation Phase Transition in Clustered Complex Networks'',
{\it Phys.\ Rev. \ X \ \bf 4}, 041020 (2014).

\bibitem{connesGRAVMAT}
A.~Connes,
``Gravity coupled with matter and the foundation of noncommutative geometry'',
{\it Commun.\ Math.\ Phys. \ \bf 183}, 1, 155--176 (1996).

\bibitem{connesRECON}
A.~Connes,
``On the spectral characterization of manifolds'',
{\it J. Noncommut.\ Geom. \ \bf 7}, 1--82 (2013).

\bibitem{masterthesis}
M. D'Arcangelo,
{\it Random Non-Commutative Geometries\/}
(Unpublished master's thesis, 2017).

\bibitem{duane}
S. Duane, A.~D. Kennedy, B.~J. Pendleton and D. Roweth,
``Hybrid Monte Carlo'',
{\it Phys.\ Lett. \ B \ \bf 195}, 216--222 (1987).

\bibitem{earl}
D.~J. Earl and M.~W. Deem,
``Parallel tempering: Theory, applications, and new perspectives'',
{\it Phys.\ Chem. \ Chem. \ Phys. \bf 7}, 3910--3916 (2005).

\bibitem{erdos}
P. Erd\H os and A. R\'enyi,
``On the Evolution of Random Graphs'',
{\it Publication of the Mathematical Institute of the Hungarian Academy of Sciences\/}
17--61 (1960).

\bibitem{glaser}
L. Glaser,
``Scaling behaviour in random non-commutative geometries'',
{\it J. Phys.\ A\ \bf 50}, 27, 275201 (2017).

\bibitem{hastings}
W.~K. Hastings,
``Monte  Carlo sampling  methods using  Markov chains  and their applications'',
{\it Biometrika\ \bf 57}, 1, 97--109 (1970).

\bibitem{montvay}
I.~Montvay and G. Munster,
{\it Quantum Fields on a Lattice\/}
(Cambridge University Press, 2009).

\bibitem{neal}
R.~M. Neal,
{\it Handbook of Markov Chain Monte Carlo\/}
(Chapman \& Hall / CRC Press, 2011).

\bibitem{suijlekom}
W.~D. Suijlekom,
{\it Noncommutative Geometry and Particle Physics\/}
(Springer, 2014).

\bibitem{swendsen}
R.~H. Swendsen and J.~-S. Wang,
``Replica Monte Carlo Simulation of Spin-Glasses'',
{\it Phys.\ Rev. \ Lett. \ \bf 57}, 1, 2607 (1986).

\bibitem{zukovic}
M. \v Zukovi\v c and G. Kalagov,
``Magnetic quasi-long-range ordering in nematic systems due to competition between higher-order couplings'',
{\it Phys.\ Rev. \ E \ \bf 97}, 052101 (2018).


\end{thebibliography}

\end{document}
